---
title: "Data Day Accompanying Script"
author: "Emily Robinson"
date: "1/23/2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = TRUE)
library(tidyverse)
library(glue)
library(magrittr)
```

## Reading in the data

We'll try the base R way first. 

```{r}
multiple_choice_responses_base <- read.csv("multipleChoiceResponses.csv")
# multiple_choice_responses_base
```

Let's say we wanted to know the numbers of NAs in each column. We can use `is.na` to change each entry in a column to TRUE or FALSE, depending on whether it's NA, and then sum the column (because `TRUE` evaluates as 1 and `FALSE` as 0) to get the total number of NAs. 

To do this for every column, we can use `purrr::map_df`, which applies the given function over the whole dataset, column by column, and returns a dataframe with the same column names and one row representing the number of NAs in each column. If you're used to the `apply` family of functions, `purrr` offers the same capabilities in a more consistent way. The `_df` part of the function represents what we want back (a dataframe); using `map` would get us a list. 

```{r}
# for one column
sum(is.na(multiple_choice_responses_base$Country))

# for every column 
multiple_choice_responses_base %>%
  purrr::map_df(~sum(is.na(.)))
```

Wow that's lucky! So many variables that don't have NAs. But ... is it too good to be true? Let's look at the entries of one column.  

```{r}
multiple_choice_responses_base %>%
  dplyr::count(StudentStatus)
```

Yep. We see here we have a lot of `""` entries instead of NAs. We can correct this with `dplyr::na_if`, which takes as an argument what we want to turn into NAs. We can also use `%<>%`, which is a reassignment pipe. 

```{r}
multiple_choice_responses_base %<>%
  dplyr::na_if("")

## is the same as: 

multiple_choice_responses_base <- multiple_choice_responses_base %>%
  na_if("")
```

Now we can count the NAs again. 

```{r}
multiple_choice_responses_base %>%
  purrr::map_df(~sum(is.na(.)))
```

And it's fixed! 

Alternative: use `readr::read_csv` instead of `read.csv`. 

```{r}
multiple_choice_responses <- readr::read_csv("multipleChoiceResponses.csv")
```

It's definitely faster, but it seems we have some errors. Let's inspect them. 

```{r}
problems(multiple_choice_responses)
```

We see the row and column where the problem occurs. What's happening is that `read_csv` uses the first 1000 rows of a column to guess its type. But in some cases, it's guessing the column is an integer, because the first 1000 rows are whole numbers, when actually it should be double, as some entries have decimal points. We can fix this by changing the number of rows `read_csv` uses to guess the column type (with the `guess_max` argument) to the number of rows in the dataset. 

```{r}
multiple_choice_responses <- readr::read_csv("multipleChoiceResponses.csv", 
                                             guess_max = nrow(multiple_choice_responses))
```

Great! 

## Initial examination 

Let's see what we can glean from the column names themselves.

```{r}
colnames(multiple_choice_responses)
```

It's clear that there were categories of questions, like "Job Factor" and "Work Methods Frequency."

Now let's take a look at our numeric columns with skimr. Skimr is a package from rOpenSci that allows you to quickly view summaries of your data. `select_if` is a great function if you want to select only columns where a certain condition is true (in this case, whether it's a numeric column).  

```{r}
multiple_choice_responses %>%
  select_if(is.numeric) %>%
  skimr::skim()
```

I love the histograms. We can quickly see from the histograms that people self teach a lot and spend a good amount of time building models and gathering data, compared to visualizing data or working in production.   

Let's see how many distinct answers we have for each question (most interesting for the non-numeric questions). `n_distinct()` is just a shorter and faster version of `length(unique())`. We can use `map_df` once again to apply a function to every column. 

```{r}
multiple_choice_responses %>%
  purrr::map_df(~n_distinct(.)) 
```

This data would be more helpful if it was tidy and had two columns, `question` and `num_distinct_answers`. We can use `tidyr::gather` to change our data from "wide" to "long" format and then `arrange` it so we can see the columns with the most distinct answers first. If you've used (or are still using!) reshape2, check out tidyr; reshape2 is retired. While not exactly equivalent, `tidyr::spread` replaces `reshape2::dcast`, `tidyr::separate` `reshape2::colsplit`, and `tidyr::gather` `reshape2::melt`. 

```{r}
multiple_choice_responses %>%
  purrr::map_df(~n_distinct(.)) %>%
  tidyr::gather(question, num_distinct_answers) %>%
  arrange(desc(num_distinct_answers))
```

Let's take a look at the question with the most distinct answers. 

```{r}
multiple_choice_responses %>%
  count(WorkMethodsSelect, sort = TRUE)
```

This is clearly multiple select question, where if a person selected multiple answers they're listed as one entry, separated by commas. Let's tidy it up. 

`!` here is short for `== FALSE`. So `!is.na(WorkMethodsSelect)` is the same as `is.na(WorkMethodsSelect) == FALSE`. `str_split` says "Take this string and split it into a list by dividing it where there are `,`s." 

```{r}
nested_workmethods <- multiple_choice_responses %>%
  select(WorkMethodsSelect) %>%
  filter(!is.na(WorkMethodsSelect)) %>%
  mutate(work_method = str_split(WorkMethodsSelect, ",")) 

nested_workmethods
```

Now we have a list column, with each entry in the list being one work method. We can unnest this so we can get back a tidy dataframe. 

```{r}
unnested_workmethods <- nested_workmethods %>%
  unnest(work_method) %>%
  select(work_method)

unnested_workmethods
```

### Graphing Frequency of Different Work Challenges 

Now I want to move on to understanding what challenges people face at work. This was one of those categories where there were multiple questions asked, all having names starting with `WorkChallengeFrequency` and ending with the challenge (e.g "DirtyData"). 

We can find the relevant columns by using the dplyr `select` helper `contains`. We then use `gather` to tidy the data for analysis, keep only the non-NAs, and remove the irrelevant part of the name for each question using `stringr::str_replace`. 

```{r}
WorkChallenges <- multiple_choice_responses %>%
  select(contains("WorkChallengeFrequency")) %>%
  gather(question, response) %>%
  filter(!is.na(response)) %>%
  mutate(question = stringr::str_replace(question, "WorkChallengeFrequency", "")) 

ggplot(WorkChallenges, aes(x = response)) + 
  geom_bar() + 
  facet_wrap(~question) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This graph has two main problems. First, there are too many histograms. But second, the order of the x-axis is wrong. We want it to go from least often to most, but instead `rarely` is in the middle. We can manually reorder the level of this variable using `forcats::fct_relevel`. 

```{r}
WorkChallenges %>%
  mutate(response = forcats::fct_relevel(response, "Rarely", "Sometimes", "Often", "Most of the time")) %>%
  ggplot(aes(x = response)) + 
  geom_bar() + 
  facet_wrap(~question) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Now we've got the x-axis in the order we want it. Let's try dichotomizing the variable by grouping "most of the time" and "often" together as the person considering something a challenge. We can use `if_else` and `%in%`. `%in%` is equivalent to `response == "Most of the time" | response == "Often"` and can save you a lot of typing if you have a bunch of variables to match. 

Grouping by the question, we can use `summarise` to reduce the dataset to one row per question, adding the variable `perc_problem` for the percentage of responses that thought something was a challenge often or most of the time. This way, we can make one graph with data for all the questions and easily compare them. 

```{r}
perc_problem_work_challenge <- WorkChallenges %>%
  mutate(response = if_else(response %in% c("Most of the time", "Often"), 1, 0)) %>%
  group_by(question) %>%
  summarise(perc_problem = mean(response)) 
```

```{r}
ggplot(perc_problem_work_challenge, aes(x = question, y = perc_problem)) + 
  geom_point() +
  coord_flip()
```

This is better, but it's hard to read because the points are scattered all over the place. Although you can spot the highest one, then you have to track it back to the correct variable. And it's also hard to tell the order of the ones in the middle. 

We can use `forcats:fct_reorder` to have the x-axis be ordered by another variable, in this case the y-axis. While we're at it, we can use `scale_y_continuous` and`scales::percent` to update our axis to display in percent and `labs` to change our axis labels. 

```{r}
ggplot(perc_problem_work_challenge, aes(x = fct_reorder(question, perc_problem), y = perc_problem)) + 
  geom_point() +
  coord_flip() + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Aspect", y = "Percentage encountering challenge frequently")
```

Much better!