---
title: "Data Day Talk Code"
author: "Emily Robinson"
date: "1/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = TRUE)
library(tidyverse)
library(glue)
library(magrittr)
```

## Reading in the data

We'll try the base R way first. 

```{r}
multiple_choice_responses_base <- read.csv("multipleChoiceResponses.csv")
multiple_choice_responses_base
```

Now we can count NAs in each column.

```{r}
# for one column
sum(is.na(multiple_choice_responses_base$Country))

# for every column 
multiple_choice_responses_base %>%
  purrr::map_df(~sum(is.na(.)))
```

No NAS - too good to be true?

```{r}
multiple_choice_responses_base %>%
  dplyr::count(StudentStatus)
```

Use `na_if` to change `""` to `NA`. 

```{r}
multiple_choice_responses_base %<>% 
  dplyr::na_if("")

## is the same as: 

multiple_choice_responses_base <- multiple_choice_responses_base %>%
  na_if("")
```

Now we can count the NAs again. 

```{r}
multiple_choice_responses_base %>%
  purrr::map_df(~sum(is.na(.)))
```

Alternative: use `readr::read_csv` instead of `read.csv`. 

```{r}
multiple_choice_responses <- readr::read_csv("multipleChoiceResponses.csv")
```

Check out the issues noted. 

```{r}
problems(multiple_choice_responses)
```

```{r}
multiple_choice_responses <- readr::read_csv("multipleChoiceResponses.csv", 
                                             guess_max = nrow(multiple_choice_responses))
```

### Examining the datset 

```{r}
colnames(multiple_choice_responses)
```

Use `skimr::skim()` to examine numeric columns with `select_if`. 

```{r}
multiple_choice_responses %>%
  select_if(is.numeric) %>%
  skimr::skim()
```

Use `n_distinct()` and `map_df` to see the number of distinct answers for every quesiton.  

```{r}
multiple_choice_responses %>%
  purrr::map_df(~n_distinct(.)) 
```

Rearrange the data into long format. 

```{r}
multiple_choice_responses %>%
  purrr::map_df(~n_distinct(.)) %>%
  tidyr::gather(question, num_distinct_answers) %>%
  arrange(desc(num_distinct_answers))
```

### Examine Work Methods

Let's take a look at one of the ones with the most distinct answers. 

```{r}
multiple_choice_responses %>%
  count(forcats::fct_infreq(WorkMethodsSelect))
```

Here people are clearly selecting multiple answers, recorded separated by commas. Let's tidy it up. 

```{r}
nested_workmethods <- multiple_choice_responses %>%
  select(WorkMethodsSelect) %>%
  filter(!is.na(WorkMethodsSelect)) %>%
  mutate(work_method = str_split(WorkMethodsSelect, ",")) 

nested_workmethods
```

We can unnest. 

```{r}
unnested_workmethods <- nested_workmethods %>%
  unnest(work_method) %>%
  select(work_method)

unnested_workmethods
```

Get the number of people using each work method. 

```{r}
method_freq <- unnested_workmethods %>%
  count(method = fct_infreq(work_method))

method_freq
```

### Examine Work Challenges

Find all columns about `WorkChallengeFrequency` using the dplyr `select` helper `contains` and clean the data. 

```{r}
WorkChallenges <- multiple_choice_responses %>%
  select(contains("WorkChallengeFrequency")) %>%
  gather(question, response) %>%
  filter(!is.na(response)) %>%
  mutate(question = stringr::str_replace(question, "WorkChallengeFrequency", "")) 

WorkChallenges
```

Try an initial plot. 

```{r}
ggplot(WorkChallenges, aes(x = response)) + 
  geom_bar() + 
  facet_wrap(~question) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Reorder the graphs. 

```{r}
WorkChallenges %>%
  mutate(response = forcats::fct_relevel(response, "Rarely", "Sometimes", "Often", "Most of the time")) %>%
  ggplot(aes(x = response)) + 
  geom_bar() + 
  facet_wrap(~question) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Dichotomize the variable to make one graph with data for all the questions and easily compare them. 

```{r}
perc_problem_work_challenge <- WorkChallenges %>%
  mutate(response = if_else(response %in% c("Most of the time", "Often"), 1, 0)) %>%
  group_by(question) %>%
  summarise(perc_problem = mean(response)) 
```

```{r}
ggplot(perc_problem_work_challenge, aes(x = question, y = perc_problem)) + 
  geom_point() +
  coord_flip()
```

Use `forcats:fct_reorder` to have the x-axis be ordered by another variable, in this case the y-axis. Use`scale_y_continuous` and`scales::percent` to u pdate our axis to display in percent and `labs` to change our axis labels. 

```{r}
ggplot(perc_problem_work_challenge, aes(x = fct_reorder(question, perc_problem), y = perc_problem)) + 
  geom_point() +
  coord_flip() + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = "Question", y = "Percentage encountering challenge frequently")
```

